Electric polarizability as the fundamental variable in the dielectric properties of two-dimensional materials

### 2018-12-17 SI.tex
Correct x-references in manuscript and SI

1. Supplementary Figure \ref{fig:S1}:  Simulations of $\varepsilon_{\mathrm{SL}}$ using many-body $G_{\rm 0}W_{\rm 0}$   \\
%
%2. Supplementary Section \ref{epsilon-model}: $\varepsilon^{\parallel, \perp}_{\mathrm{SL}}$ dependence on $L$ \\ 
%
2. Supplementary Section \ref{sec:2D-3D-rescale}: Hypothetical 2D ``dielectric constant'' rescaled from 3D dielectric constant \\
%
3. Supplementary Section \ref{ssec:theory-1}: Polarizability-Based Theoretical Model  \\ 
%
4. Supplementary Section \ref{sec:pol-2D-Eg}: Dependence of $\alpha_{\mathrm{2D}}$ on bandgap \\ 
%
5.

@@

\section{Further analysis on the dielectric properties of 2D materials}

@@

\subsection{Profile of induced dipoles of 2D material}
\label{sec:dipole-plot}
Here we show in detail the
$\Delta \rho=\rho(\boldsymbol{E}) - \rho(\boldsymbol{E}=0)$ profile of
the 2H-MoS$_{2}$ slab in main text Figure 1a. The
density $\rho$ is calculated as the average value in the
\textit{xy}-plane, and the value is shown in Figure
\ref{fig:rho-profile}. As can be seen the induced charges on the
MoS$_{2}$ layer only extends to a width of $\sim{}$12 \AA{}. When the
SL size $L \gg$12 \AA{} as in the first principle calculations done in
the main text, the induced dipoles from the periodic images do not
interact, and thus giving converged values of $\alpha_{\mathrm{2D}}$.

\begin{figure}[htbp]
 \centering
 \includegraphics[width=0.5\linewidth]{img/SI-dipole-MoS2.pdf}
 \caption{$\Delta \rho$ as a function of $z$ around the
   MoS$_{2}$ layer, corresponding to main text Figure
   \ref{main-fig-1}a. Green and red parts corresponding to negative and
   positive induced charges. The external electric field
   $E_{\mathrm{ext}}$ is 0.01 eV/\AA{}.}
 \label{fig:rho-profile}
\end{figure}

@@

We calculate the frequency-dependent dielectric properties of hBN with
varying supperlattice $L$ at PBE, G$_{0}$W$_{0}$, and G$_{0}$W$_{0}$
with electron-hole interaction (G$_{0}$W$_{0}$+BSE). As shown in
Figure \ref{fig:PBE-omega-in} and \ref{fig:PBE-omega-out}, the
intensities of $\varepsilon_{\mathrm{SL}}^{\parallel}(\omega)$ and
$\varepsilon_{\mathrm{SL}}^{\perp}(\omega)$ reduce with increasing $L$
throughout the whole frequency range. The $L$-dependency can also be
removed using Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-ref}, yielding lattice-independent
polarizabilities throughout the frequency domain.

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-PBE-omega-in_plane.pdf}
 \caption{Dependence of in-plane dielectric properties on $L$ for
   monolayer BN calculated using PBE method: \textbf{a} Real part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{b} Real part of
   $\alpha_{\mathrm{2D}}^{\parallel}$, \textbf{c} Imaginary part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{d} Imaginary part of
   $\alpha_{\mathrm{2D}}^{\parallel}$.  The same {\bf k}-sampling is used
   in all simulations with the only variable quantity being $L$. }
  \label{fig:PBE-omega-in}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\linewidth]{img/SI-PBE-omega-out_of_plane.pdf}
   \caption{Dependence of out-of-plane dielectric properties on $L$ for
   monolayer BN calculated using PBE method: \textbf{a} Real part of
   $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{b} Real part of
   $\alpha_{\mathrm{2D}}^{\perp}$, \textbf{c} Imaginary part of
   $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{d} Imaginary part of
   $\alpha_{\mathrm{2D}}^{\perp}$.  The same {\bf k}-sampling is used
   in all simulations with the only variable quantity being $L$. }
  \label{fig:PBE-omega-out}
\end{figure}

@@

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-omega-in_plane.pdf}
  \caption{Similar to Figure \ref{fig:PBE-omega-in} but calculated using G$_{0}$W$_{0}$ method. Peak shift can be observed in the dielectric function spectra.
}
  \label{fig:GW-omega-in}
\end{figure}

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-omega-out_of_plane.pdf}
 \caption{Similar to Figure \ref{fig:PBE-omega-out} but calculated
   using G$_{0}$W$_{0}$ method. Peak shift can be observed in the
   dielectric function spectra.}
  \label{fig:GW-omega-out}
\end{figure}

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-BSE-in_plane.pdf}
  \caption{Similar as in Fig. \ref{fig:GW-omega-in} but taking into account electron-hole interactions at the level of 
  the Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ - BSE). 
  The peak shift in energies are reduced in comparison with G$_{\rm 0}$W$_{0}$.
}
  \label{fig:BSE-omega-in}
\end{figure}

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-BSE-out_of_plane.pdf}
  \caption{Similar as in Fig. \ref{fig:GW-omega-out} but taking into account electron-hole interactions at the level of 
  the Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ - BSE). 
  The peak shift in energies are reduced in comparison with G$_{\rm 0}$W$_{0}$, HSE06 and PBE but still present. 
}
  \label{fig:BSE-omega-out}
\end{figure}

@@

%%%%%%\section{Theoretical Analysis}
%%%%%%\label{sec:theory}
\section{Polarizability-Based Theoretical Model}
\label{ssec:theory-1}
The universal relations for $\alpha_{\mathrm{2D}}^{\parallel}$ and
$\alpha_{\mathrm{2D}}^{\perp}$ revealed by main text
Eqs. \ref{main-eq:2D-Moss-para} and \ref{main-eq:2D-Moss-perp} are not
coincidence. Combining recent theoretical findings of the linear
relation between exciton binding energy $E_{\mathrm{b}}$ and
$E_{\mathrm{g}}$ of 2D materials
\cite{Choi_linear_2015,Olsen_2016_hydrogen,Jiang_2017_Eg_Eb}, and the
fact that the $E_{\mathrm{b}}$ is roughly inversely proportional to
$\alpha_{\mathrm{2D}}^{\parallel}$ \cite{Pulci_2014}, it is reasonable
to have a general Moss-like relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{-1}$. Moreover,
the bandgap-independent relation of 2D $\alpha_{\mathrm{2D}}^{\perp}$
resembles molecular polarizabilities of conjugate molecules
\cite{Davies_1952}, fullerenes \cite{Sabirov_2014} and carbon
nanotubes \cite{Benedict_1995}, which are also shown to be
geometry-dependent.

In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wave function of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components, similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$, such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, c, v are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{eq:alpha-para-def} and \ref{eq:alpha-perp-def}.

{\bf ES:} There is something weird with the cross-referencing for the equations 3a and 3b between the main text and 
SI. They changed to 2a and 2b in the SI.

\subsection{Detailed derivations for main text Eq. \ref{eq:alpha-para-def} 3a}  %
\label{sssec:theory-1}

In this section we show how main text Eq. \ref{eq:alpha-para-def}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{v}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Present the summation of k-space into integral, and further combining
with the Bloch wave representation and the
$\mathbf{k} \cdot \mathbf{p}$ theory\cite{Jiang_2017_Eg_Eb}, we arrive
at Eq. \ref{eq:alpha-para-def}. Note that from
Ref. \citenum{Jiang_2017_Eg_Eb}, the coefficient between
$\alpha^{\parallel}_{\mathrm{2D}}$ and $E_{\mathrm{g}}$ is given by:
$\alpha_{\mathrm{2D}}^{\parallel} = N e^{2}/(2\pi E_{\mathrm{g}})$,
where $N$ is the degeneracy associated with the minimal bandgap. Such
relation predicts the linear correlation coefficient between
$(4\pi \varepsilon_{0})/\alpha^{\parallel}_{\mathrm{2D}}$ (measured in
\AA{}$^{-1}$) and $E_{\mathrm{g}}$ (measured in eV) as
$8 \pi^{2} \varepsilon_{0} \AA{} / (e N) \approx 0.436 / N$. From Figure 2{\bf c}, we have
$(4 \pi \varepsilon_{0}\cdot \text{\AA{}})/\alpha_{\mathrm{2D}}^{\parallel}
\sim{} 0.183 E_{\mathrm{g}}/ \mathrm{eV}$, corresponding to $N$
between 2 and 3, which is a reasonable result for the 2D materials
studied.

{\bf ES:} Tian, please include more details in this section on the derivations!

\subsection{Detailed derivations for main text Eq. \ref{eq:alpha-perp-def}}
\label{ssec:theory-2}

For main text Eq. \ref{eq:alpha-perp-def}, treating the in-plane
wave functions as plane wave with form
$\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the matrix
element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z-2}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at:
\begin{equation}
\label{eq:alpha_perp_RPA}
\alpha_{\mathrm{2D}}^{\perp} = \frac{2e^{2}}{(2 \pi) ^{2}} \int d^{2}\mathbf{k}
\sum_{\mathrm{c, v}}
\frac{|<\psi_{\mathrm{v}}(\mathbf{k})|z|\psi_{\mathrm{c}}(\mathbf{k})>|^{2}}
{E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
The formalism is slightly different from Eq.\ref{eq:alpha_para_RPA}.

>>>>

\begin{itemize}
\item Supplementary Section \ref{sec:polarizability-analysis}: Further analysis on the dielectric properties of 2D materials

\item Supplementary Section \ref{sec:2D-3D-rescale}: Hypothetical 2D ``dielectric constant'' rescaled from 3D dielectric constant

\item Supplementary Section \ref{sec:theory-1}: Polarizability-Based Theoretical Model

\item Supplementary Section \ref{sec:pol-2D-Eg}: Dependence of $\alpha_{\mathrm{2D}}$ on bandgap

\item Supplementary Section \ref{sec:gpaw}: Validation on the C2DB 2D materials database

\item Supplementary Section \ref{sec:2D-3D}: More discussion about the
  relation between 2D and 3D properties

  \item Supplementary Section \ref{sec:aniso}: More discussions about the dielectric anisotropy

\item Supplementary Section \ref{sec:raw}: Raw data from first
  principles calculations

\end{itemize}

@@

\section{Further analysis on the dielectric properties of 2D materials}
\label{sec:polarizability-analysis}

@@

\subsection{Profile of induced dipoles of 2D material}
\label{sec:dipole-plot}
Here we show in detail the
$\Delta \rho=\rho(\boldsymbol{E}) - \rho(\boldsymbol{E}=0)$ profile of
the 2H-MoS$_{2}$ slab in main text Figure \ref{main-fig-1}. The
density $\rho$ is calculated as the average value in the
\textit{xy}-plane, and the value is shown in Figure
\ref{fig:rho-profile}. As can be seen the induced charges on the
MoS$_{2}$ layer only extends to a width of $\sim{}$12 \AA{}. When the
SL size $L \gg$12 \AA{} as in the first principle calculations done in
the main text, the induced dipoles from the periodic images do not
interact, and thus giving converged values of $\alpha_{\mathrm{2D}}$.

\begin{figure}[htbp]
 \centering
 \includegraphics[width=0.5\linewidth]{img/SI-dipole-MoS2.pdf}
 \caption{$\Delta \rho$ as a function of $z$ around the
   MoS$_{2}$ layer, corresponding to main text Figure
   \ref{main-fig-1}\textbf{a}. Green and red parts corresponding to negative and
   positive induced charges. The external electric field
   $E_{\mathrm{ext}}$ is 0.01 eV/\AA{}.}
 \label{fig:rho-profile}
\end{figure}

@@

We calculate the frequency-dependent dielectric properties of hBN with
varying supperlattice $L$ at PBE, G$_{0}$W$_{0}$, and G$_{0}$W$_{0}$
with electron-hole interaction (G$_{0}$W$_{0}$+BSE). As shown in
Figure \ref{fig:PBE-omega-in} and \ref{fig:PBE-omega-out}, the
intensities of $\varepsilon_{\mathrm{SL}}^{\parallel}(\omega)$ and
$\varepsilon_{\mathrm{SL}}^{\perp}(\omega)$ reduce with increasing $L$
throughout the whole frequency range. The $L$-dependency can also be
removed using Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-def}, yielding lattice-independent
polarizabilities throughout the frequency domain.

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-PBE-omega-in_plane.pdf}
 \caption{Dependence of in-plane dielectric properties on $L$ for
   monolayer BN calculated using PBE method: \textbf{a} Real part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{b} Real part of
   $\alpha_{\mathrm{2D}}^{\parallel}$, \textbf{c} Imaginary part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{d} Imaginary part
   of $\alpha_{\mathrm{2D}}^{\parallel}$.  The same {\bf k}-sampling
   is used in all simulations with the only variable quantity being
   $L$. Clearly, the $L$-dependency of the superlattice dielectric
   function is removed using 2D polarizabilities.}
  \label{fig:PBE-omega-in}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.9\linewidth]{img/SI-PBE-omega-out_of_plane.pdf}
  \caption{Dependence of out-of-plane dielectric properties on $L$ for
    monolayer BN calculated using PBE method: \textbf{a} Real part of
    $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{b} Real part of
    $\alpha_{\mathrm{2D}}^{\perp}$, \textbf{c} Imaginary part of
    $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{d} Imaginary part of
    $\alpha_{\mathrm{2D}}^{\perp}$.  The same {\bf k}-sampling is used
    in all simulations with the only variable quantity being
    $L$. Clearly, the $L$-dependency of the superlattice dielectric
    function is removed using 2D polarizabilities.}
  \label{fig:PBE-omega-out}
\end{figure}

@@

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-omega-in_plane.pdf}
 \caption{Similar to Figure \ref{fig:PBE-omega-in} but calculated
   using G$_{0}$W$_{0}$ method. A shift of peak position (frequency)
   in the dielectric function spectra is observed, resulted from the
   change of quasi-particle energy in G$_{0}$W$_{0}$ calculations
   invloving the 2D slab.}
  \label{fig:GW-omega-in}
\end{figure}

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-omega-out_of_plane.pdf}
 \caption{Similar to Figure \ref{fig:PBE-omega-out} but calculated
   using G$_{0}$W$_{0}$ method.  A shift of peak position (frequency)
   in the dielectric function spectra is observed, resulted from the
   change of quasi-particle energy in G$_{0}$W$_{0}$ calculations
   invloving the 2D slab.}
  \label{fig:GW-omega-out}
\end{figure}

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-BSE-in_plane.pdf}
 \caption{Similar as in Fig. \ref{fig:GW-omega-in} but taking into
   account electron-hole interactions at the level of the
   Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ - BSE).  The peak shift
   in energies are reduced in comparison with G$_{\rm 0}$W$_{0}$. As a
   result the polarizabilities becomes almost independent of $L$,
   consistent to results obtained by the PBE method.}
  \label{fig:BSE-omega-in}
\end{figure}

\begin{figure}[H]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-BSE-out_of_plane.pdf}
 \caption{Similar as in Fig. \ref{fig:GW-omega-out} but taking into
   account electron-hole interactions at the level of the
   Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ - BSE).  As a result
   the polarizabilities becomes almost independent of $L$, consistent
   to results obtained by the PBE method.  }
  \label{fig:BSE-omega-out}
\end{figure}

@@

%%%%%%\section{Theoretical Analysis}
%%%%%%\label{sec:theory}
\section{Polarizability-Based Theoretical Model}
\label{sec:theory-1}
The universal relations for $\alpha_{\mathrm{2D}}^{\parallel}$ and
$\alpha_{\mathrm{2D}}^{\perp}$ revealed by main text
Eqs. \ref{main-eq:2D-Moss-para} and \ref{main-eq:2D-Moss-perp} are not
coincidence. Combining recent theoretical findings of the linear
relation between exciton binding energy $E_{\mathrm{b}}$ and
$E_{\mathrm{g}}$ of 2D materials
\cite{Choi_linear_2015,Olsen_2016_hydrogen,Jiang_2017_Eg_Eb}, and the
fact that the $E_{\mathrm{b}}$ is roughly inversely proportional to
$\alpha_{\mathrm{2D}}^{\parallel}$ \cite{Pulci_2014}, it is reasonable
to have a general Moss-like relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{-1}$. Moreover,
the bandgap-independent relation of 2D $\alpha_{\mathrm{2D}}^{\perp}$
resembles molecular polarizabilities of conjugate molecules
\cite{Davies_1952}, fullerenes \cite{Sabirov_2014} and carbon
nanotubes \cite{Benedict_1995}, which are also shown to be
geometry-dependent.

In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wave function of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components, similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$, such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, c, v are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{main-eq:2D-Moss-para} and
\ref{main-eq:2D-Moss-perp}.

% {\bf ES:} There is something weird with the cross-referencing for the equations 3a and 3b between the main text and 
% SI. They changed to 2a and 2b in the SI.

\subsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-para}}  %
\label{ssec:theory-1-para}

In this section we show how main text Eq. \ref{main-eq:alpha-para-def}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Following the method of $\mathbf{k} \cdot \mathbf{p}$ theory from
Ref. \citenum{Jiang_2017_Eg_Eb}, the matrix element in the numerator
of Eq. \ref{eq:alpha_para_RPA} is approximated by:

\begin{equation}
\label{eq:matrix-approx}
|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}
\approx {\displaystyle \frac{\hbar^{2}}{2 m^{*}}
  \frac{1}{E_{\mathrm{g}} + \frac{\hbar^{2} k^{2}}{2 m^{*}}}}
\end{equation}
plug it into Eq. \ref{eq:alpha_para_RPA} and integrate within the 2D Brillouin Zone from $|k|=0$ to $|k|= k_{\mathrm{BZ}}$, where $k_{\mathrm{BZ}}$ is the wavevector at the boundary of the 2D Brillouin Zone, we get:
\begin{equation}
  \label{eq:derivation-2D-Moss-para}
  \begin{aligned}
\alpha_{\mathrm{2D}}^{\parallel} &= N\cdot -\frac{e^{2}}{2 \pi}
\frac{1}{E_{g} + \beta} \Biggr\vert_{\beta=0}^{\beta=\frac{\hbar^{2} k^{2}_{\mathrm{BZ}}}{m^{*}}} \\
&\approx N e^{2}/(2\pi E_{\mathrm{g}}) = C^{\parallel} E_{g}^{-1}
\end{aligned}
\end{equation}
where $N$ is de degeneracy of bands associated with $E_{g}$. The
approximation in Eq. \ref{eq:derivation-2D-Moss-para} is due to the
fact that $\frac{\hbar^{2} k^{2}_{\mathrm{BZ}}}{m^{*}} \gg E_{g}$, and
we arrive at main text Eq. \ref{main-eq:2D-Moss-para}.

The coefficient of $C^{\parallel}$ adapted from
Ref. \citenum{Jiang_2017_Eg_Eb} $C^{\parallel} = N e^{2}/(2\pi)$
predicts linear correlation between $\alpha^{\parallel}_{\mathrm{2D}}$
and $E_{\mathrm{g}}^{-1}$. We validate this by examining the
DFT-calculated
$(4\pi \varepsilon_{0})/\alpha^{\parallel}_{\mathrm{2D}}$ (measured in
\AA{}$^{-1}$) and $E_{\mathrm{g}}$ (measured in eV) in main text
Figure \ref{main-fig-3}\textbf{c}. The coefficient $C^{\parallel}$
becomes
$8 \pi^{2} \varepsilon_{0} \AA{} / (e N) \approx 0.436 / N = 0.183$,
corresponding to $N$ between 2 and 3, which is a reasonable result for
the 2D materials studied.

% {\bf ES:} Tian, please include more details in this section on the derivations!

\subsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-perp}}
\label{ssec:theory-1-perp}
